---
title: "Tidyverse Assignment"
author: "Alain T Kuiete"
date: "12/5/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How accurated is the FiveThirtyEigth Model on Predicting Soccer Match Scores

### Library
```{r}
library(tidyverse)
```

### Loading the Datasets
```{r}
soccer <- read.csv("https://raw.githubusercontent.com/AlainKuiete/DATA607/master/spi_matches.csv")
```

### Introspecting the dataset
```{r}
head(soccer)
summary(soccer)
str(soccer)
```

### Selecting Useful Variables Using Pipe and Predicting Score 
```{r}
pred.res <- soccer %>% select(spi1, spi2, score1, score2) %>%                       mutate(sc.p=if_else(spi1>=spi2, 1, 0))
head(pred.res)
```

### Actual scores
```{r}
act.pred <- pred.res %>% mutate(sc.r=if_else(score1>=score2, 1, 0))%>% select(sc.r, sc.p)
head(act.pred)
```

```{r}
act.pred <- act.pred %>% mutate(diff=if_else(sc.r==sc.p, 1, 0))
```


```{r}
res <- act.pred %>% group_by(diff)%>% summarise(count=n())
res
```

### Accuracy of prediction

```{r}
pt <- as.numeric(filter(res, count, diff==1)[1,2])
pf <- as.numeric(filter(res, count, diff==0)[1,2])
```

```{r}
pt
pf
ac <- pt/(pt+pf) 
ac
```

### Overall, FiveThirtyEigth predicted at 60.6% the 2016 Europeean  Leagues of Soccer.



### Classification of teams by leagues in 2016 World Soccer leagues

```{r}
s.teams <- select(soccer, league_id, league, team1, team2, score1, score2)
```

### we can imbricate if else statements

```{r}
s.teams <- s.teams %>% group_by(league_id, league, team1)%>% mutate(pt1=if_else(score1>score2,3, if_else(score1==score2, 1, 0))) %>% mutate(pt2=if_else(score2>score1,3, if_else(score2==score1, 1, 0)))
```

### Points gained by each team
```{r}
 s.teams <- s.teams %>% summarise_at( c("pt1","pt2"), sum, na.rm = TRUE) %>% mutate(pts = pt1+pt2)
```


### Best teams by leagues in 2016 Wold Soccer leagues

```{r}
s.teams %>% filter( pts==max(pts))
```

